Predicting Which States are Most Interested in the term *China Virus* Based off Identity Politics, Demography and COVID-19 Statistics 

### 2020-05-04 Final-Draft.Rmd
visuals

```{r, warning=FALSE, include=FALSE, message=FALSE}
library(dplyr)
library(usmap)
library(maps)
library(ggplot2)
library(tidyverse)
library(ggmap)
library(viridis)
library(rgdal)
library(gridExtra)
library(plotly)
library(janitor)
library(reshape2)
library(tidyr)
library(rstan)
library(rstanarm)
library(bayesplot)
library(sf)
library(remotes)
```

@@

```{r}
Finaldata$hover <- with(Finaldata, paste(polyname, "<br>","Positive Cases:", positive, "<br>","Negative Cases:", negative,"<br>",
                           "Total Cases:", totalTestResults,"<br>", "% White:", percent_white,"<br>",
                            "% Asian:", percent_asian, "<br>", "State Color:", StateColor,
                        "<br>", "Total Population", total_population))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

@@

fig
```
```{r}
Finaldata$hover <- with(Finaldata, paste(polyname, "<br>","Positive Cases:", positive, "<br>","Negative Cases:", negative,"<br>",
                           "Total Cases:", totalTestResults,"<br>", "% White:", percent_white,"<br>",
                            "% Asian:", percent_asian, "<br>", "State Color:", StateColor,
                        "<br>", "Total Population", total_population))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

fig <- plot_geo(Finaldata, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~percent_white, text = ~hover, locations = ~State,
    color = ~percent_white, colors = 'Blues'
  )
fig <- fig %>% colorbar(title = "Interest")
fig <- fig %>% layout(
    title = '2020 Percent White by State <br>(Hover for additional information)', geo=g
  )

fig
```
```{r}

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}
Trump<- as.Date("03/16/2020", format = "%m/%d/%Y")
day1<-as.Date("03/10/2020",format = "%m/%d/%Y")
day2<-as.Date("03/24/2020",format = "%m/%d/%Y")

google_ts<-google%>%
  mutate(Day = as.Date(as.character(Day)))

a<-google_ts %>% 
  #filter(Day<=day2)%>%
  #filter(Day>=day1)%>%
  group_by(Region, Day) %>% 
  summarize(ChinaVirusSearch = median(ChinaVirusInterest)) %>% 
  ggplot(aes(x=Day, y=ChinaVirusSearch, color=Region))+
  geom_point()

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}
c <-  ggplot(Finaldata, aes(y=positive, x=Winner, fill=(Winner)))+
      geom_boxplot() +
      facet_wrap(~Region)+
      theme(legend.position = "none") +
      scale_fill_manual(values = c("Democrat"="blue3", "Republican"= "red3"))+
      lims(y= c(0,1000))+
      ggtitle("Positive COVID-19 Cases by State Political Party Winner & Region")
```

```{r, warning=FALSE}
ggplotly(c)
```
This visualisation depicts the distribution of positive COVID-19 cases by region and by which political party won in the 2016 elections. We can see that Democrat states in the Midwest, Mountain, and West have a larger range and higher quantile metrics for positive cases overall. For the Northeast and South regions the mean of positive COVID-19 cases are higher but not significantly. This is an interesting pattern considering that poltical party affiliation appears to interact with the number of postive cases by region.

@@

## Model 1 Repeated Measures Model

  For our simplest model we decided to use a repeated measures model. Our team decided that the repeated measures model was necessary component because of how our data is set up. As we can see in our dataset, each state has a value for their `ChinaVirusInterest` for each day in our target period (2020-03-14 - 2020-03-21). Given the ability to use repeated measures and our prior understanding of the varying characteristics (demographic,political,covid-impact) within different states, the repeated measures model allows us to capture these differences in `ChinaVirusInterest` with the $\theta_i$ value which represents each state's mean value.

### Model Structure

$$\begin{aligned}
Y_{ij}|\theta_i, \mu, \sigma_w, \sigma_b \sim N(\theta_i,\sigma_w^2)\\
\theta_i|\mu,\sigma_b \overset{ind}{\sim} N(\mu, \sigma_b^2)\\
\sigma_b,\sigma_w \sim Exp(...)
\end{aligned}$$

$Y_{ij} = $ `ChinaVirusInterest` per $i=State$, and $j= Day$
$\theta_i = $ State i's unique mean value
$\sigma_w = $ within state variation
$\sigma_b = $ between state variation

```{r}
model_data<-
  Finaldata%>%
  mutate(Day=as.numeric(Day)-as.numeric(min(Day)))%>%
  select(ChinaVirusInterest, Day, percent_white, StateColor, positive, State)

model_data<- na.omit(model_data)
```

```{r}
#Repeated Measures Model
set.seed(454)
RM_model <- stan_glmer(
  ChinaVirusInterest ~ (1 | State),
  data = model_data, family = gaussian,
)
```

### 3.2 Model Descriptions

@@

# 4. Model Evaluation (RESULTS)

## 4.1 Model 1 Evaluation (Repeated measures)

>>>>

```{r, warning=FALSE, include=FALSE, message=FALSE}
library(dplyr)
library(ggrepel)
library(usmap)
library(maps)
library(ggplot2)
library(tidyverse)
library(ggmap)
library(viridis)
library(rgdal)
library(gridExtra)
library(plotly)
library(janitor)
library(reshape2)
library(tidyr)
library(rstan)
library(rstanarm)
library(bayesplot)
library(sf)
library(remotes)
```

@@

```{r, echo=FALSE, warning=FALSE}
Finaldata$hover <- with(Finaldata, paste(polyname, "<br>","Positive Cases:", positive, "<br>","Negative Cases:", negative,"<br>",
                           "Total Cases:", totalTestResults,"<br>", "% White:", percent_white,"<br>",
                            "% Asian:", percent_asian, "<br>", "State Color:", StateColor,
                        "<br>", "Total Population", total_population))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

@@

#########white fig
whitefig <- plot_geo(Finaldata, locationmode = 'USA-states')
whitefig <- whitefig %>% add_trace(
    z = ~percent_white, locations = ~State,
    color = ~percent_white, colors = 'Blues'
  )
whitefig <- whitefig %>% colorbar(title = "% White")
whitefig <- whitefig %>% layout(
    title = 'Percent White Identifying by State', geo=g
  )

#### Asian fig
asianfig <- plot_geo(Finaldata, locationmode = 'USA-states')
asianfig <- asianfig %>% add_trace(
    z = ~percent_asian, locations = ~State,
    color = ~percent_asian, colors = 'Greens'
  )
asianfig <- asianfig %>% colorbar(title = "% Asian")
asianfig <- asianfig %>% layout(
    title = 'Percent Asian Identifying by State', geo=g
  )

fig

whitefig

asianfig

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}

a<-Finaldata %>% 
  #filter(Day<=day2)%>%
  #filter(Day>=day1)%>%
  group_by(Region, Day) %>% 
  summarize(ChinaVirusSearch = median(ChinaVirusInterest)) %>% 
  ggplot(aes(x=Day, y=ChinaVirusSearch, color=Region))+
  geom_point()+ geom_line()

@@

```{r, warning=FALSE, message= FALSE, echo=FALSE}
c <-  ggplot(Finaldata, aes(y=positive, x=Winner, fill=(Winner)))+
      geom_boxplot() +
      facet_wrap(~Region)+
      theme(legend.position = "none") +
      scale_fill_manual(values = c("Democrat"="blue3", "Republican"= "red3"))+
      lims(y= c(0,1000))+
      ggtitle("Positive COVID-19 Cases by State Political Party Winner & Region")
ggplotly(c)
```
This visualisation depicts the distribution of positive COVID-19 cases by region and by which political party won in the 2016 elections. We can see that Democrat states in the Midwest, Mountain, and West have a larger range and higher quantile metrics for positive cases overall. For the Northeast and South regions the mean of positive COVID-19 cases are higher but not significantly. This is an interesting pattern considering that poltical party affiliation appears to interact with the number of postive cases by region.

@@

## 3.1 Markov Chain Algorithm

## 3.2 Models of Interest

## 3.3 Model Descriptions

@@

# 4. Model Evaluation

## 4.1 Model 1 Evaluation( Repeated measures)
